---
title: "Homework_2"
author:
  - Håkon Sandaker
  - Vincent Wilmet
date: "3/25/2021"
output: pdf_document
---

# Comparison of estimators
## a) MLE with expectation  and variance

We have the Binomial Distribution
$$
Bin(n, \theta) \sim \binom n k \theta^k (1-\theta)^{n-k}
$$

Log-Likelhood:
$$
\begin{aligned}
\text{log-likelihood} &= log(Bin(n, \theta)) \\
&= log(\binom n k \theta^k (1-\theta)^{n-k}) \\
&= log(\binom n k) + k * log(\theta) + (n-k) * log(1-\theta)
\end{aligned}
$$
MLE:
$$
\begin{aligned}
\frac{\partial \text{ log-likelihood}}{\partial \theta} &= 0 \\
\frac{\partial log(\binom n k) + k * log(\theta) + (n-k) * log(1-\theta)}{\partial \theta} &= 0 \\
\frac{k}{\theta} - \frac{n-k}{1-\theta} &= 0 \\
\frac{k}{\theta} &= \frac{n-k}{1-\theta} \\
\frac{1-\theta}{\theta} &= \frac{n-k}{k} \\
\frac{1}{\theta} - 1 &= \frac{n}{k} - 1 \\
\frac{1}{\theta} &= \frac{n}{k} \\
\theta &= \frac{k}{n} \\
\hat{\theta}_{MLE} &= \frac{k}{n}
\end{aligned}
$$
Expectation and Variance:
$$
E[\hat{\theta}_{MLE}] = E[\frac{k}{n}] = \frac{1}{n}E[k] = \frac{\theta n}{n} = \theta
$$
$$
\begin{aligned}
Var(\hat{\theta}_{MLE}) &= Var(\frac{k}{n}) \\
&= \frac{1}{n^2} Var(k)  \text{  (by variance property)}\\
&= \frac{1}{n^2} n \theta (1 - \theta) \\
&= \frac{\theta (1 - \theta)}{n}
\end{aligned}
$$

## b)
$$
\hat{\theta}_{alt.} = \frac{X+1}{n+2}
$$
$$
\begin{aligned}
E[\hat{\theta}_{alt.}] &= E[\frac{X+1}{n + 2}] \\
&= \frac{E[X] + 1}{n+2} \text{  (by linearity of expectation)}\\
&= \frac{n \theta + 1}{n + 2}
\end{aligned}
$$

$$
\begin{aligned}
Var(\hat{\theta}_{alt.}) &= Var(\frac{X+1}{n + 2}) \\
&= \frac{1}{(n + 2)^2} Var(X + 1) \\
&= \frac{Var(X)}{(n+2)^2} \\
&= \frac{n \theta (1-\theta)}{(n+2)^2}
\end{aligned}
$$

## c) MSE
### MLE
$$
\begin{aligned}
MSE(\hat{\theta}_{MLE}) &= (E[\hat{\theta}_{MLE}] - \theta)^2 + Var(\hat{\theta}_{MLE}) \\
&= (\theta - \theta)^2 + \frac{\theta(1-\theta)}{n} \\
&= \frac{\theta(1-\theta)}{n}
\end{aligned}
$$

### Alt
$$
\begin{aligned}
MSE(\hat{\theta}_{alt.}) &= (E[\hat{\theta}_{alt.}] - \theta)^2 + Var(\hat{\theta}_{alt.}) \\
&= (\frac{n \theta + 1}{n + 2} - \theta)^2 + \frac{n \theta (1-\theta)}{(n+2)^2} \\
&= (\frac{n \theta + 1}{n + 2} - \frac{\theta (n + 2)}{n + 2})^2 + \frac{n \theta (1-\theta)}{(n+2)^2} \\
&= (\frac{n \theta + 1}{n + 2} - \frac{n \theta + 2 \theta)}{n + 2})^2 + \frac{n \theta (1-\theta)}{(n+2)^2} \\
&= (\frac{1 - 2 \theta}{n + 2})^2 + \frac{n \theta (1-\theta)}{(n+2)^2} \\
&= \frac{(1 - 2 \theta)^2}{(n + 2)^2} + \frac{n \theta (1-\theta)}{(n+2)^2} \\
&= \frac{(1 - 2 \theta)^2 + n \theta (1-\theta)}{(n + 2)^2} \\
&= \frac{(1 - 2 \theta)^2 + n \theta (1-\theta)}{(n + 2)^2} \\
&= \frac{1 - 4 \theta + 4 \theta^2 + n \theta - n \theta^2}{(n + 2)^2} \\
&= \frac{1 - \theta (n - 4) (\theta - 1)}{(n + 2)^2} \\
\end{aligned}
$$
### MSE vs Alt.
The maximum likelihood estimator of $\hat{\theta}_{MLE}$ is unbiased as it is equal to its true variance. Hence, we would rather use the MSE MLE over the MSE Alt. approach.

## d) Comparisson
```{r}
n <- 10
theta <- seq(0, 1, 0.01)

MSE_MLE <- function(n, theta)
{
  return (theta*(1 - theta)/n)
}

MSE_Alt <- function(n, theta)
{
  return ((1 - theta*(n-4)*(theta-1))/ (n+2)^2)
}

# Plot the MLE & Alt
plot(theta, MSE_MLE(n, theta), col="blue", main="Mean Squared Error", ylab="MSE")
lines(theta, MSE_Alt(n, theta), col="red", main="Mean Squared Error", ylab="MSE")
legend("topright", c("MLE","Alt."), fill=c("blue", "red"))
```

# 2. Robustness of the estimators
## a)
Find the MLE of the Guassian Distribution.
Likelihood:
$$
\begin{aligned}
\text{likelihood} &= \prod_n \frac{1}{\sigma \sqrt{2 \pi}}  e^{- \frac{1}{2}(\frac{x - \theta}{\sigma})^2} \\
&= \prod_n \frac{1}{\sqrt{2 \pi}}  e^{- \frac{1}{2}(x - \theta)^2}, \text{from the fact that } \sigma \text{ is 1} \\
&= \frac{1}{(2 \pi)^{\frac{n}{2}}} e^{-\frac{1}{2} \sum_{i=1}^n (x_i - \theta)^2}
\end{aligned}
$$
Log-Likelihood:
$$
\begin{aligned}
\text{log-likelihood} &= log(\frac{1}{(2 \pi)^{\frac{n}{2}}} e^{-\frac{1}{2} \sum_{i=1}^n (x_i - \theta)^2}) \\
&= -\frac{n}{2} log(2 \pi) - \frac{1}{2} \sum_{i=1}^n (x_i - \theta)^2 \\
\end{aligned}
$$

MLE:
$$
\begin{aligned}
\frac{\partial \text{ log-likelihood}}{\partial \theta} &= 0 \\
\frac{\partial -\frac{n}{2} log(2 \pi) - \frac{1}{2} \sum_{i=1}^n (x_i - \theta)^2}{\partial \theta} &= 0 \\
\sum_{i=1}^n (x_i - \theta) &= 0 \\
\sum_{i=1}^n x_i - n \theta &= 0 \\
\theta = \frac{\sum_{i=1}^n x_i}{n} = \bar{x}
\end{aligned}
$$

Thus,
$\hat{Q}^{MV} \sim N(\bar{X}_n, 1)$

## b)
Of course, the data might not follow the Gaussian distribution. If thats the case, the MLE will not work. However, if the size of sample n is sufficiently large, by CLT we can say that the distribution will converge towards a Gaussian distribution, at which point $\hat{Q}^{MV}$ is an appropriate estimator.

## c)
Expectation:
$$
E[Q] = 0.99 * E[P_0] + 0.01 * E[P_{300}] = 0 + 0.01 * 300 = 3
$$

Variance:
$$
Var(Q) = Var(0.99 * P_0) + Var(0.01 * P_{300}) = 0.99^2 * Var(P_0) + 0.01^2 * Var(P_{300}) = 0.99^2 + 0.01^2 = 0.9802.
$$

Since the variance of Q is not equal to 1 it does not belong to the model $N(\theta, 1)$. On the contrary, Q is a mixture of the two distributions.

Density of Q:
$$
f_x(x) = \frac{1}{\sqrt{2 \pi}} e ^{- \frac{1}{2} (x - 3)^2}
$$

## d)
Yes, but $\hat{Q}^{MV}$ is not identical to Q. $\hat{Q}^{MV} \sim  N(\bar{X}_n, 1) \sim N(3, 1)$, but $Q \sim N(3, 0.99)$ with a different variance.

## e)
We are given
$$
\Phi^{-1}(\frac{1}{2 * 0.99}) = 0.013
$$
Additionally, we have
$$
Q = 0.99P_{0} + 0.01P_{300}
$$

This tells us that 99% of the mixed distribution is weighted towards the $P_0$ distribution. We have that $P_0 \sim N(0, 1)$. Furthermore, we have the property that the median of the standard normal distribution is the same as the mean. Hence the median of the quantile $\Phi^{-1}(\frac{1}{2 * 0.99}) = 0.013$ is supposed to be close to zero, which it is.

# 3. Hypothesis testing and doping controls
## a)
Hematocrit levels in the blood:
$$Y \sim N(45, 2)$$

Observed values:
$$X \sim N(\mu_1, 2)$$
$$X \sim N(\mu_2, 2)$$

Want to check if expectation of X is equal or greater than the mean of 45.

## b)
Hypothesis.
$$
\begin{aligned}
H_0: \mu_i &= 45 \\
H_1: \mu_i &> 45 \\
\end{aligned}
$$

## c)
Estimator.
$$
Z = \frac{\bar{x}_i - 45}{\sqrt{2}}, \space Z \sim N(0, 1)
$$
With $H_0$ Z follows N(0, 1).
With $H_1$ Z does not follow N(0, 1).

## d)
```{r}
qnorm(.95)
```

With $p-value=.05$ and a right tail we get a Z-score of 1.64.
Hence, we reject the null hypothesis if Z value is greater than 1.64.

## e)
### i.
<!-- Type 1 Error -->
```{r}
pnorm(45, mean=45, sd=2, lower.tail=FALSE)
```

### ii.
```{r}
pnorm(60, mean=45, sd=2, lower.tail=FALSE)
```
Very unlikely.

### iii.
```{r}
cv <- qnorm(.95, mean=45, sd=2)
cv
```

## f)
### i.
Reject if Z is above 1.645.

### ii.
J.C: $Z-score = \frac{48-45}{\sqrt{2}} = 2.121$. Reject.
S.R: $Z-score = \frac{50-45}{\sqrt{2}} = 3.53$. Reject.

### iii. using the student population t test
```{r}
typeI.test <- function(mu0, sigma, n, alpha, iterations = 10000) {
  pvals <- rep(NA, iterations)
  for(i in 1 : iterations){
    temporary.sample <- rnorm(n = n, mean = mu0, sd = sigma)
    temporary.mean <- mean(temporary.sample)
    temporary.sd <- sd(temporary.sample)
    pvals[i] <- 1 - pt((temporary.mean - mu0)/(temporary.sd / sqrt(n)), df = n-1)
  }
  return(mean(pvals < alpha))
}

typeI.test(mu0 = 45, sigma = 2, n = 5, alpha = 0.05)
```

## g)
### i.
```{r}
theta <- seq(30, 60, 0.5)
# changed from mean=45 to critical value=48.28971
plot(theta, 1 - pnorm(theta, mean=cv, sd=2), col="blue", main="Power function", xlab="theta", ylab="1 - beta(theta)")
```

### ii.
By looking at the Power function, the probability of detecting an abnormal hematocrit level is around 19%.
```{r}
1 - pnorm(50, mean=cv, sd=2)
```


# 4.
## a)
```{r}
df <- read.csv("poulpeF.csv")
head(df)
```

## b) 
Mean and Variance:
```{r}
mean <- mean(df$Poids)
variance <- var(df$Poids)
mean
variance
```

Under the MLE:
```{r}
# MASS library with fitdistr
library(ggplot2)
library(MASS)
library(stats4)
MLE <- fitdistr(as.numeric(df$Poids), "normal")
MLE$estimate["mean"]
MLE$estimate["sd"]^2
```

## c)
```{r}
hist(df$Poids, main="Histogram")
```
No, it is skewed towards the right. Hence, basing it on a Gaussian can be a problem.

## d)
95% Confidence interval
$$
\bar{X} \pm Z * \frac{std}{\sqrt{n}}
$$

```{r}
err <- qnorm(.975) * sqrt(variance) / sqrt(length(df$Poids))
# left, lower bound
mean - err
# right, upper bound
mean + err
```
